---
title: "Assignment 4"
author: "Madison Herrboldt"
date: "2022-12-02"
output: html_document
---
## Working with dplyr

```{r}
#load required library
library(tidyverse)

#load in dataset
Matt_ebird <- read_csv("https://github.com/mbtoomey/Biol_7263/blob/main/Data/MBT_ebird.csv?raw=true")
view(Matt_ebird)
```

#### Question 1
In which year did I observe the most individual birds? How many?
2014: 9,303 birds
```{r}
year_count <- Matt_ebird %>% #created new variable that will be just the year and counts
  group_by(year) %>% #group the data by year
  summarize(total_birds = sum(count)) #get the sum of the counts for each year
view(year_count)
#arrange into descending order to make it easier to see what year had the most birds
arrange(year_count, by = desc(total_birds))
```

#### Question 2
In that year how many different species of birds did I observe?
210
```{r}
species_2014 <- Matt_ebird %>% #create new variable
  filter(year == 2014) %>% #filter by only the year 2014
  count(scientific_name) #list each species and their observed counts
view(species_2014)
nrow(species_2014) #output the number of rows (number of different species observed)
```

#### Question 3 
In which state did I most frequently observe Red-winged Blackbirds?
Missouri
```{r}
RWBB <- Matt_ebird %>% #create new variable
  filter(common_name == "Red-winged Blackbird") %>% #filter by common name
  count(location, sort = TRUE) #count how many birds were seen in each location and then sort by high to low
head(RWBB)
```

#### Question 4
Filter observations for a duration between 5 and 200 minutes. Calculate the mean rate per checklist that I encounter species each year. Specifically, calculate the number of species in each checklist divided by duration and then take the mean for the year.
```{r}
duration_filter <- Matt_ebird %>%
  filter(duration >= 5 & duration <= 200) %>% #filter by duration to be equal to/between 5 and 200
  group_by(list_ID) %>% #group by checklist
  summarise(duration = first(duration), year = first(year), count = n_distinct(common_name)) #summarize duration and species count per checklist and also include the year, not sure if my count function is actually doing what I want here
rate_birds <- mutate(duration_filter, rate = (count/duration)) %>% #calculate the rate per checklist
  group_by(year) %>% #group by year
  summarise(mean_rate = mean(rate)) #calculate the mean rate per year
head(rate_birds)
```

#### Question 5
Create a tibble that includes the complete observations for the top 10 most frequently observed species. First generate a top 10 list and then use this list to filter all observations. Export this tibble as a .csv file saved to a folder called “Results” folder within your R project and add link to the markdown document.
```{r}
bird_freq <- Matt_ebird %>%  
  group_by(scientific_name) %>%  #group by species
  mutate(frequency = sum(count)) %>% #calculate frequency per species
  slice(which.max(frequency)) #index by most frequently seen

Top_10_data <- head(arrange(bird_freq, by = desc(frequency)),10) #create new table arranged by the top ten most 
#frequently seen bird in descending order of frequency

write_csv(Top_10_data, "Results/Top_10_data.csv")
```

[Top 10 Data](Results/Top_10_data.csv)



